---
title: 'DS202: Homework 5'
author: "Muhammad Khairi Norizan"
date: "11/9/2021"
output: html_document
---

## Data Engineering
```{r, warning=FALSE, message=FALSE}
# import library 
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)

# import data set
dataframe <- read_csv("2021_Iowa_Liquor_Sales.csv") 

# data cleaning
dataframe$Date <- mdy(dataframe$Date)
dataframe <- dataframe %>% 
              separate(`Store Location`, c('POINT', 'Longitude', 'Latitude'), " ", convert = TRUE) %>% 
              separate(`Longitude`, c('SymbolLon','Longitude'), sep ='\\(') %>% 
              separate(`Latitude`, c('Latitude','SymbolLat'), sep ='\\)') %>% 
              select(-c('POINT', 'SymbolLon', 'SymbolLat'))
```


## Question 1
First overview: Provide a visual breakdown of the liquor category (by `Category Name`). Include volume sold in the breakdown. Describe your figure. Which type of liquor is the most popular?

```{r, message=FALSE}
categoryBreakdown <- dataframe %>% 
                      group_by(`Category Name`) %>% 
                      summarize(VolumeSold = sum(`Volume Sold (Gallons)`, na.rm = TRUE)) %>% 
                      arrange(desc(VolumeSold))

print(categoryBreakdown)
```

```{r}
top10Category <- categoryBreakdown %>% head(10)

ggplot(top10Category, aes(x=`Category Name`, y=VolumeSold, fill = `Category Name`)) + 
  geom_bar(stat = "identity") + 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90, hjust=1))
```
\

## Question 2
Which day of the week has the most orders? Use a barchart to visualize the results. Explain the pattern and provide a reason why the pattern occurs.

```{r}
dfWithWeekday <- dataframe %>% mutate(Weekday = weekdays(Date))

weekdayBreakdown <- dfWithWeekday %>% 
                      group_by(Weekday) %>% 
                      count(Weekday, sort = TRUE)

ggplot(weekdayBreakdown, aes(x=Weekday, y=n, fill = Weekday)) + 
  geom_bar(stat = "identity") + 
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust=1)) + 
  labs(y = "Count")
```

## Question 3
Create a single plot for the monthly volumes sold. Each liquor category should be shown using a separate time series. Label the top 5 categories using text labels.

```{r}
dfWithMonth <- dataframe %>% mutate(Month = month(Date))

categoryMonthBreakdown <- dfWithMonth %>% 
                          group_by(`Category Name`, Month) %>% 
                          count(`Category Name`, Month, sort = TRUE)

top5Category <- categoryBreakdown %>% head(5)

top5MonthlyBreakdown <- categoryMonthBreakdown %>% 
                        filter(`Category Name` %in% top5Category$`Category Name`) %>% 
                        arrange(desc(`Category Name`))

ggplot(top5MonthlyBreakdown, aes(x=Month, y=n)) +
  geom_line() + 
  facet_wrap(~`Category Name`) + 
  labs(y = "Count")
```

## Question 4
Which categories are spring graduation party favorites? Find at least one category.

```{r}
dfWithSemester <- dataframe %>% 
                  mutate(Semester = semester(Date)) %>% 
                  mutate(Semester = ifelse(Semester == 1, "Spring", "Fall"))

categorySemesterBreakdown <- dfWithSemester %>% 
                              filter(Semester == 'Spring') %>% 
                              group_by(`Category Name`) %>% 
                              summarize(BottlesSold = sum(`Bottles Sold`, na.rm = TRUE))

top10CategoryBySemester <- categorySemesterBreakdown %>% head(10)

ggplot(top10CategoryBySemester, aes(x=`Category Name`, y=BottlesSold, fill = `Category Name`)) +
  geom_bar(stat = "identity") + 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90, hjust=1))
```